---
title: "Artist Similarity"
output: html_notebook
---

```{r}
load("large data/Lastfm_data.RData")
require(lsa)
require(psych)
require(NMF)
require(VennDiagram)
```

Examine the 0.05th quantiles of the number of unique artists listened to by users.  This information can be helpful in subsetting users that should be used in the training utility matrix.
```{r}
quantile(User.plays$N.artists, seq(0, 1, .05), na.rm = T)
```

User's with a more diverse listening history should allow for better recommendations (less sparsity in utility matrix).  Let's make recommendations off of the listening history of users who are in the top 20% in terms of number of unique artists.  If you would like to allow for a larger or smaller subset of users, feel free to change the minium number of unique artists, but don't save/overwrite the altered script.
```{r}
User.plays.50 <- User.plays[which(User.plays$N.artists >= 40),]
```

Similarly to how we have eliminated users with rather specific musical preference (as inidicated by a low number of unique artists listened to), It is valuable to eliminate artists who have a narrow listener following (as indicated by a low number of unique listeners).  Let's examine the 0.05th quantiles of Artist.plays$N.users to get a feel the user listening base of many artists.
```{r}
quantile(Artist.plays$N.users, seq(0.9, 1, .01))
```

Let's only consider artists who are in the top 4% by number of unique users (160+ unique listeners, as a arbitrarily round number).  This should cut down on the number of recommendations of niche/specific artists.  This number can be changed moving forward, but for the sake of developing a functional algorithm, let's stick to artists who have a somewhat broad following.  This still allows for close to 10,000 artists, so it's not an overly restrictive library of artists.
```{r}
Artist.plays.100 <- Artist.plays[which(Artist.plays$N.users >= 500),]
```

Similarly to to "Music Recommendation Simulation_v1.Rmd", we must create a data sample from which to train the recommender.  One modification to this sample is that it should include users who are randomly sampled from "User.plays.50" AND only artists who appear in "Artist.plays.10".
```{r}
set.seed(04072017)
N_samp <- 1000
samp.users <- User.plays.50$User[sample(c(1:length(User.plays.50$User)), N_samp , replace = F)]
samp.dat <- Lastfm[which(Lastfm$usersha1 %in% samp.users & Lastfm$artname %in% unique(Artist.plays.100$Artist)),]
```

Let's see how many unique artists there are in our sample of N users
```{r}
print(paste("Number of unique artists in a sample of", N_samp,"users:", 
        length(unique(samp.dat$artname))))
```

As in "Music Recommendation Simulation_v1", we must create a utility matrix of users' musical preferences.  The rows of the matrix are each of the unique users in our sample, the columns are each of the unique artists in our sample.  The values of each cell are the number of listens by user[i] of artist[j].  This matrix is sparse, as users don't listen to most of the different artists, and artists aren't listened to by most of the users.
```{r}
samp.Utility <- matrix(data = NA, nrow = length(unique(samp.dat$usersha1)),
                       ncol = length(unique(samp.dat$artname)))
rownames(samp.Utility) <- unique(samp.dat$usersha1)
colnames(samp.Utility) <- unique(samp.dat$artname)

# fill in utility matrix for training.
for (i in 1:nrow(samp.dat)) {
  samp.Utility[which(rownames(samp.Utility) == samp.dat$usersha1[i]), which(colnames(samp.Utility) == samp.dat$artname[i])] <- 1
}
rm(i)
samp.Utility[which(is.na(samp.Utility))] <- 0
```

Create matrix of artist similarity by cosine distance
```{r}
artist.similarity <- cosine(samp.Utility)
```

Input an artist name, and generate the top N similar artists (by cosine distance).
```{r}
artist.name <- "the beatles"
N <- 10

similar.artists <- artist.similarity[which(colnames(artist.similarity) == artist.name),]
similar.artists <- similar.artists[order(similar.artists, decreasing = T)][2:(N+1)]

similar.artists.df <- data.frame(artist = names(similar.artists), similarity = as.numeric(similar.artists))

print(paste(sum(samp.Utility[,which(colnames(samp.Utility) == artist.name)]), " different users listen to ", artist.name, sep = ""))
similar.artists.df

```

Create venn diagrams of artist
```{r}
venn.df <- matrix(c(samp.Utility[,which(colnames(samp.Utility) == artist.name)], samp.Utility[,which(colnames(samp.Utility) == similar.artists.df$artist[1])], samp.Utility[,which(colnames(samp.Utility) == similar.artists.df$artist[2])]), ncol = 3, byrow = F)

listeners.1 <- max(length(venn.df[which(venn.df[,1] == 1)]), 0)
listeners.2 <- max(length(venn.df[which(venn.df[,2] == 1)]), 0)
listeners.3 <- max(length(venn.df[which(venn.df[,3] == 1)]), 0)
listeners.12 <- max(length(venn.df[which(venn.df[,1] == 1 & venn.df[,2] == 1)]), 0)
listeners.13 <- max(length(venn.df[which(venn.df[,1] == 1 & venn.df[,3] == 1)]), 0)
listeners.23 <- max(length(venn.df[which(venn.df[,2] == 1 & venn.df[,3] == 1)]), 0)
listeners.123 <- max(length(venn.df[which(venn.df[,1] == 1 & venn.df[,2] == 1 & venn.df[,3] == 1)]),0)

# Venn Diagram
grid.newpage()
draw.triple.venn(area1 = listeners.1, area2 = listeners.2, area3 = listeners.3, n12 = listeners.12, n23 = listeners.23, n13 = listeners.13, n123 = listeners.123, category = c(artist.name, as.character(similar.artists.df$artist[1]), as.character(similar.artists.df$artist[2])), lty = "blank", fill = c("skyblue", "pink1", "mediumorchid"), scaled = T)


```

